---
title: 'PRECAST: Human Breast Cancer Data Analysis'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRECAST: Human Breast Cancer Data Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE,
  fig.width = 11,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
```

The package can be loaded with the command:
```{r  eval = TRUE}
library("PRECAST")
```


## Fit PRECAST using real data: 2 sliced Human breast cancer

Loading data
```{r  eval=FALSE}
dir.file <- "E:\\Research paper\\IntegrateDRcluster\\iDR-SC\\Human_Breast_Cancer\\BlockA_Section"
seuList <- list()
for(r in 1:2){
  message("r = ", r)
  seuList[[r]] <- read10XVisium(paste0(dir.file, r))
}
library(Seurat)
seuList
## Create 
library(PRECAST)
PRECASTObj <-  CreatePRECASTObject(seuList)
human_breast_cancer2 <- PRECASTObj
save(human_breast_cancer2, file= paste0(dir.file, 1, "\\human_breast_cancer2_PRECASTObj.rds"))
```

### load the  Human breast cancer data and Prepare the PRECASTObject.

Human breast cancer data have been pre-processed and saved as the `PRECASTObj` format.
```{r  eval = TRUE}
## Create 
dir.file <- "E:\\Research paper\\IntegrateDRcluster\\iDR-SC\\Human_Breast_Cancer\\BlockA_Section1\\"
load(paste0(dir.file, "human_breast_cancer2_PRECASTObj.rds"))
PRECASTObj <- human_breast_cancer2
rm(human_breast_cancer2)
```
### Add the model setting
```{r  eval = TRUE}
## check the number of genes/features after filtering step
PRECASTObj@seulist

## Add adjacency matrix list for a PRECASTObj object to prepare for PRECAST model fitting.
PRECASTObj <-  AddAdjList(PRECASTObj, platform = "Visium")

## Add a model setting in advance for a PRECASTObj object. verbose =TRUE helps outputing the information in the algorithm.
PRECASTObj <- AddParSetting(PRECASTObj, Sigma_equal=FALSE, verbose=TRUE, int.model=NULL)

```


### Fit PRECAST 
For function `PRECAST`, users can specify the number of clusters $K$ or set `K` to be an integer vector by using modified BIC(MBIC) to determine $K$. First, we try using user-specified number of clusters. For convenience, we give the selected number of clusters by MBIC (K=14).
```{r  eval = TRUE}
### Given K
# PRECASTObj <- PRECAST(PRECASTObj, K=14)

# To save compilation time ,we use the cached data.
# save(PRECASTObj, file= paste0(dir.file,  "\\human_breast_resList_PRECASTObj.rds"))
load(paste0(dir.file,  "\\human_breast_resList_PRECASTObj.rds"))
```

Select a best model 
```{r  eval = TRUE}
## backup the fitting results in resList
resList <- PRECASTObj@resList
PRECASTObj <- selectModel(PRECASTObj)

```

Integrate the two samples by the function `IntegrateSpaData`.
```{r  eval = TRUE}
seuInt <- IntegrateSpaData(PRECASTObj, species='Human')
seuInt 
## The low-dimensional embeddings obtained by PRECAST are saved in PRECAST reduction slot.
```
Show the spatial scatter plot for clusters
```{r  eval = TRUE, fig.height = 4, fig.width=7.5}
cols_cluster <-  c( "#FD7446" ,"#709AE1", "#31A354","#9EDAE5",
 "#DE9ED6" ,"#BCBD22", "#CE6DBD" ,"#DADAEB" ,
 "yellow", "#FF9896","#91D1C2", "#C7E9C0" ,"#6B6ECF", "#7B4173" )
p12 <- SpaPlot(seuInt, batch=NULL,point_size=1, cols=cols_cluster, combine=TRUE)
p12
# users can plot each sample by setting combine=FALSE
```

Show the spatial UMAP/tNSE RGB plot to illustrate the performance in extracting features.

```{r  eval = TRUE, fig.height = 3, fig.width=5}
seuInt <- AddUMAP(seuInt) 
p13 <- SpaPlot(seuInt, batch=NULL,item='RGB_UMAP',point_size=2, combine=TRUE, text_size=15)
p13
#seuInt <- AddTSNE(seuInt) 
#SpaPlot(seuInt, batch=NULL,item='RGB_TSNE',point_size=2, combine=T, text_size=15)
```


Show the tSNE plot based on the extracted features from PRECAST to check the performance of integration.
```{r  eval = TRUE}
seuInt <- AddTSNE(seuInt, n_comp = 2) 
library(patchwork)
p1 <- dimPlot(seuInt, item='cluster', point_size = 0.5, font_family='serif', cols=cols_cluster) # Times New Roman
p2 <- dimPlot(seuInt, item='batch', point_size = 0.5,  font_family='serif')
p1 + p2 
```


Combined differential expression analysis 
```{r  eval = TRUE}
library(Seurat)
dat_deg <- FindAllMarkers(seuInt)
library(dplyr)
n <- 10
dat_deg %>%
  group_by(cluster) %>%
  top_n(n = n, wt = avg_log2FC) -> top10

seuInt <- ScaleData(seuInt)
seus <- subset(seuInt, downsample = 400)


```

Plot DE genes' heatmap for each spatial domains identified by PRECAST.
```{r  eval = TRUE, fig.height = 9, fig.width=11}
color_id <- as.numeric(levels(Idents(seus)))
library(ggplot2)
## HeatMap
p1 <- doHeatmap(seus, features = top10$gene, cell_label= "Domain",
                grp_label = F, grp_color = cols_cluster[color_id],
                pt_size=6,slot = 'scale.data') +
  theme(legend.text = element_text(size=16),
        legend.title = element_text(size=18, face='bold'),
        axis.text.y = element_text(size=7, face= "italic", family='serif'))
p1
```

Save some data
```{r  eval=FALSE}
save(seuInt, file= paste0(dir.file, 1, "\\human_breast_cancer2_seuInt.rds"))
save(dat_deg, file= paste0(dir.file, 1, "\\human_breast_cancer2_DEGs.rds"))
```

## Session information
```{r  eval = TRUE}
sessionInfo()
```