---
title: 'PRECAST: DLPFC Single Sample Analysis'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRECAST: DLPFC Single Sample Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE,
  fig.width = 11,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
```
This vignette introduces the PRECAST workflow for the analysis of single spatial transcriptomics dataset. The workflow consists of three steps

* Independent preprocessing and model setting
* Probabilistic embedding and clustering  using PRECAST model
* Downstream analysis (i.e. visualization of clusters and embeddings)

We demonstrate the use of PRECAST to one human dorsolateral prefrontal cortex Visium data that are in the data R package `DataPRECAST`, which can be installed by the following command:
```{r  eval =FALSE}
remotes::install_github("feiyoung/DataPRECAST")
```

This data is also available at Github website:

* https://github.com/feiyoung/DR-SC.Analysis/tree/main/data/DLPFC_data/


The package can be loaded with the command:
```{r  eval = TRUE}
library(PRECAST)
library(DataPRECAST)
library(Seurat)
```


## Compare PRECAST with DR-SC in analyzing one sample

Loading data  from `DataPRECAST` package
```{r  eval = TRUE}
data("dlpfc_151672")
dlpfc_151672 ## a list including two Seurat object
```

### Prepare the PRECASTObject.

Create a PRECASTObj object to prepare for PRECAST models.
```{r  eval = TRUE}
library(PRECAST)
preobj <- CreatePRECASTObject(seuList = list(dlpfc_151672))
```
### Add the model setting
```{r  eval = TRUE}
## check the number of genes/features after filtering step
preobj@seulist
## Add adjacency matrix list for a PRECASTObj object to prepare for PRECAST model fitting.
PRECASTObj <-  AddAdjList(preobj, platform = "Visium")
## Add a model setting in advance for a PRECASTObj object. verbose =TRUE helps outputing the
## information in the algorithm.
PRECASTObj <- AddParSetting(PRECASTObj, Sigma_equal = FALSE, coreNum = 1,
                            maxIter=30, verbose = TRUE)

```


### Fit PRECAST 
For function `PRECAST`, users can specify the number of clusters $K$ or set `K` to be an integer vector by using modified BIC(MBIC) to determine $K$. Here, we use user-specified number of clusters. 
```{r  eval = TRUE}
### Given K
PRECASTObj <- PRECAST(PRECASTObj, K= 7)
```

Select a best model 
```{r  eval = TRUE}
## backup the fitting results in resList
resList <- PRECASTObj@resList
PRECASTObj <- selectModel(PRECASTObj)
ari_precast <- mclust::adjustedRandIndex(PRECASTObj@resList$cluster[[1]], PRECASTObj@seulist[[1]]$layer_guess_reordered)
```

Integrate the reults into a Seurat object seuInt by the function `IntegrateSpaData`.
```{r  eval = TRUE}
seuInt <- IntegrateSpaData(PRECASTObj, species='Human')
seuInt 
## The low-dimensional embeddings obtained by PRECAST are saved in PRECAST reduction slot.
```
Save the spatial and tSNE scatter plots for clusters  from PRECAST
```{r  eval = TRUE}
p_sp1 <- SpaPlot(seuInt, item='cluster', point_size = 3, combine = F)[[1]] + cowplot::theme_cowplot() + 
  ggplot2::ggtitle(paste0("PRECAST: ARI=", round(ari_precast, 2)) ) +
  ggplot2::xlab("row") + ggplot2::ylab("col")
seuInt <- AddTSNE(seuInt,n_comp = 2)
p_tsne <- dimPlot(seuInt, item='cluster')
p_tsne <- p_tsne + cowplot::theme_cowplot() + ggplot2::ggtitle("PRECAST")
```



Fit DR-SC and Plot the spatial and tSNE scatter plots for clusters 
```{r  eval = TRUE}
seu_drsc <- DR.SC::DR.SC(PRECASTObj@seulist[[1]], K=7, verbose=T)
ari_drsc <- mclust::adjustedRandIndex(seu_drsc$spatial.drsc.cluster, PRECASTObj@seulist[[1]]$layer_guess_reordered)
p_tsne_drsc <- DR.SC::drscPlot(seu_drsc)
p_tsne_drsc <- p_tsne_drsc + ggplot2::ggtitle("DR-SC")
p_sp2 <- DR.SC::spatialPlotClusters(seu_drsc)+ cowplot::theme_cowplot()  + 
  ggplot2::ggtitle(paste0("DR-SC ARI=", round(ari_drsc, 2)) ) 
```

Compare the clustering performance of PRECAST and DR-SC.

```{r  eval = TRUE}
library(patchwork)
p_sp1 + p_sp2

```


Compare the tSNE visualiztion performance of PRECAST and DR-SC.
```{r  eval = TRUE}
p_tsne + p_tsne_drsc
```

## Session information
```{r  eval = TRUE}
sessionInfo()
```