---
title: 'PRECAST: simulation'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRECAST: simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Generate the simulated data
First, we generate the multiple spatial transcriptomics data with lattice neighborhood, i.e. ST platform by using the function `gendata_seulist` in `PRECAST` package.
```{r  eval = FALSE}
library(Seurat)
library(PRECAST)
seuList <- gendata_seulist(height1=20, width1=20, height2=24, width2=25,p=400, K=4, alpha=10, sigma2=0.1)
seuList
head(seuList[[1]]) 
## Must include the columns named "row" and "col" for saving the spatial coordinates

```

## Fit PRECAST using simulated data

### Prepare the PRECASTObject with preprocessing step.
```{r  eval = FALSE}
## Create 
PRECASTObj <-  CreatePRECASTObject(seuList)

```
### Add the model setting
```{r  eval = FALSE}
## check the number of genes/features after filtering step
PRECASTObj@seulist

## Add adjacency matrix list for a PRECASTObj object to prepare for PRECAST model fitting.
PRECASTObj <-  AddAdjList(PRECASTObj, platform = "ST")

## Add a model setting in advance for a PRECASTObj object. verbose =TRUE helps outputing the information in the algorithm.
PRECASTObj <- AddParSetting(PRECASTObj, Sigma_equal=FALSE, maxIter=30, verbose=TRUE)
```

### Fit PRECAST 
For function `PRECAST`, users can specify the number of clusters $K$ or set `K` to be an integer vector by using modified BIC(MBIC) to determine $K$. First, we try using user-specified number of clusters. For convenience, we give a single K here.
```{r  eval = FALSE}
### Given K
PRECASTObj <- PRECAST(PRECASTObj, K=4)

```

Select a best model and use ARI to check the performance of clustering
```{r  eval = FALSE}
## backup the fitting results in resList
resList <- PRECASTObj@resList
# PRECASTObj@resList <- resList
PRECASTObj <- selectModel(PRECASTObj)
true_cluster <- lapply(seuList, function(x) x$true_cluster)
str(true_cluster)
mclust::adjustedRandIndex(unlist(PRECASTObj@resList$cluster), unlist(true_cluster))
```

Integrate the two samples by the function `IntegrateSpaData`.
```{r  eval = FALSE}
seuInt <- IntegrateSpaData(PRECASTObj, species='unknown')
seuInt 
## The low-dimensional embeddings obtained by PRECAST are saved in PRECAST reduction slot.
```
Show the spatial scatter plot for clusters
```{r  eval = FALSE}
p12 <- SpaPlot(seuInt, batch=NULL,point_size=2, combine=TRUE)
p12
# users can plot each sample by setting combine=FALSE
```

Show the spatial UMAP/tNSE RGB plot

```{r  eval = FALSE}
seuInt <- AddUMAP(seuInt) 
SpaPlot(seuInt, batch=NULL,item='RGB_UMAP',point_size=2, combine=TRUE, text_size=15)

#seuInt <- AddTSNE(seuInt) 
#SpaPlot(seuInt, batch=NULL,item='RGB_TSNE',point_size=2, combine=T, text_size=15)
```

Show the tSNE plot based on the extracted features from PRECAST to check the performance of integration.
```{r  eval = FALSE}
seuInt <- AddTSNE(seuInt, n_comp = 2) 
library(patchwork)
cols_cluster <- c("#E04D50", "#4374A5", "#F08A21","#2AB673", "#FCDDDE",  "#70B5B0", "#DFE0EE" ,"#D0B14C")
p1 <- dimPlot(seuInt,  font_family='serif', cols=cols_cluster) # Times New Roman
p2 <- dimPlot(seuInt, item='batch', point_size = 1,  font_family='serif')
p1 + p2 
# It is noted that only sample batch 1 has cluster 4, and only sample batch 2 has cluster 7. 
```
Show the UMAP plot based on the extracted features from PRECAST.
```{r  eval = FALSE}
dimPlot(seuInt, reduction = 'UMAP3', item='cluster', cols=cols_cluster, font_family='serif')
```

Users can also use the visualization functions in Seurat package:
```{r  eval = FALSE}
DimPlot(seuInt, reduction = 'position')
DimPlot(seuInt, reduction = 'tSNE')
DimPlot(seuInt, reduction = 'PRECAST')
```




Combined differential expression analysis 
```{r  eval = FALSE}
dat_deg <- FindAllMarkers(seuInt)
library(dplyr)
n <- 10
dat_deg %>%
  group_by(cluster) %>%
  top_n(n = n, wt = avg_log2FC) -> top10

seuInt <- ScaleData(seuInt)
seus <- subset(seuInt, downsample = 400)
color_id <- as.numeric(levels(Idents(seus)))

library(ggplot2)

## HeatMap
p1 <- doHeatmap(seus, features = top10$gene, cell_label= "Domain",
                grp_label = F, grp_color = cols_cluster,
                pt_size=6,slot = 'scale.data') +
  theme(legend.text = element_text(size=16),
        legend.title = element_text(size=18, face='bold'),
        axis.text.y = element_text(size=7, face= "italic", family='serif'))
p1

```







## Session information
```{r  eval = FALSE}
sessionInfo()
```
